---
title: "Thesis Boneyard - Unused Scripts"
author: "Katie Harrington"
date: "May 13, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



# Boneyard

## Unused state analyses and plots


## Q2c: Is there a seasonal difference in mean daytime bout lengths? Nighttime bout lengths? Are winter nights more fitful?
This calculates daytime state bout length (seconds) for the pooled active state and again for the unpooled active states (low and high activity). This creates stateBout, stateBoutSummary, state3boutDayIndMeans (seconds), and state4boutDayIndMeans. 


```{r warning=FALSE}
# Need test for samples of unequal lengths (summer days are longer)
# Results: Mean bout lengths have a exponential/Poisson/Gamma distribution? Cannot use Shapiro test for normality because n>10000.

stateBout <- data_states %>% 
  group_by(season,ID) %>% 
  arrange(dttz) %>%
  mutate(boutNum = cumsum(replace_na(lag(state_classif) != state_classif, 1))) %>%
  ungroup %>% 
  group_by(season,ID,boutNum) %>% 
  mutate(boutLengthSec = n()) %>% #bout length is in seconds
  ungroup
stateBout$state_classif <- as.factor(stateBout$state_classif)
stateBout$state <- ifelse(stateBout$state_classif==1|stateBout$state_classif==2,"rest","active")
stateBout$hr <- hour(stateBout$dttz)
stateBout$hr <- as.factor(stateBout$hr)

# Create SolarMidnight based hourly bins
stateBout$solarMidnight <- hms::as.hms((stateBout$solarnoon- dhours(12)),tz = tzOffset)
stateBout$time <- hms::as.hms(stateBout$dttz,tz = tzOffset)
stateBout$timeBin <- floor(as.numeric(difftime(stateBout$time, stateBout$solarMidnight,units = "hours")))
stateBout$timeBin <- ifelse(stateBout$timeBin < 0,stateBout$timeBin+24,stateBout$timeBin)
stateBout$timeBin <- as.factor(stateBout$timeBin)

# Calculate bout length
stateBoutSummary <- stateBout %>%
  group_by(season,ID,boutNum,state_classif,dayNight,timeBin,state) %>%
  summarize(boutLengthSec=first(boutLengthSec)) %>%
  ungroup
stateBoutSummary
rm(stateBout)

# Determine probability distributions of bout lengths

# Test for normality of summer daytime state 3 mean bout lengths (seconds)
hist(stateBoutSummary$boutLengthSec[stateBoutSummary$season=="Summer" & 
                                           stateBoutSummary$dayNight=="day" & 
                                           stateBoutSummary$state_classif=="3"]) 
# Test for normality of summer daytime state 4 mean bout lengths (seconds)
hist(stateBoutSummary$boutLengthSec[stateBoutSummary$season=="Summer" & 
                                           stateBoutSummary$dayNight=="day" & 
                                           stateBoutSummary$state_classif=="4"])

# Calculate mean state bout for state 3 per indidivual
state3boutDayIndMeans <- stateBoutSummary %>% 
  group_by(season,ID,dayNight,state_classif,boutLengthSec) %>%
  filter(dayNight=="day",
         state_classif=="3") %>%
  ungroup %>% 
  group_by(ID,season) %>%
  summarize(meanBoutLengthSec=mean(boutLengthSec))

# Calculate mean state bout for state 4 per indidivual
state4boutDayIndMeans <- stateBoutSummary %>% 
  group_by(season,ID,dayNight,state_classif,boutLengthSec) %>%
  filter(dayNight=="day",
         state_classif=="4") %>%
  ungroup %>% 
  group_by(ID,season) %>%
  summarize(meanBoutLengthSec=mean(boutLengthSec))

shapiro.test(state4boutDayIndMeans$meanBoutLengthSec)
hist(state4boutDayIndMeans$meanBoutLengthSec)
```

###Plot of seasonal daytime mean bout lengths per state
This plots seasonal mean state bouts (active and rest) by solar hour. 
KJH: Add plotline of astronomical sunrise and sunset. Look at FrontiersMovementAnalysis.

```{r warning=FALSE}
# Plot summer mean state bouts (active and rest) by solar hour. 
PboutSU <- stateBoutSummary %>% 
  group_by(season,ID,dayNight,timeBin,state) %>%
  summarize(mboutLengthMin=(mean(boutLengthSec)/60)) %>%
  ungroup %>% # Do I need to ungroup and regroup?
  group_by(season,ID,dayNight,timeBin,state,mboutLengthMin) %>%
  filter(season=="Summer") %>%
  ggplot() +
  geom_boxplot(aes(x = timeBin, y = mboutLengthMin, 
                   color = state)) +
  scale_x_discrete(breaks = seq(0, 24, by = 1)) +
  labs(x="Hour of Day",
       y = "Bout Length (min)",
       title = "Summer mean activity bout by solar hour") + 
  theme_classic(base_size = 13) + 
  theme(plot.title = element_text(size=14)) +
  theme(axis.text = element_text(size = 14, color = "black"),
        axis.title.x = element_text(color="black", size=14),
        axis.title.y = element_text(color="black", size=14))
PboutSU

# Plot winter mean state bouts (active and rest) by solar hour. 
PboutWI <- stateBoutSummary %>% 
  group_by(season,ID,dayNight,timeBin,state) %>%
  summarize(mboutLengthMin=(mean(boutLengthSec)/60)) %>%
  ungroup %>% # Do I need to ungroup and regroup?
  group_by(season,ID,dayNight,timeBin,state,mboutLengthMin) %>%
  filter(season=="Winter") %>%
  ggplot() +
  geom_boxplot(aes(x = timeBin, y = mboutLengthMin, 
                   color = state)) +
  labs(x="Hour of Day",
       y = "Bout Length (min)",
       title = "Winter mean activity bout by solar hour") + 
  theme_classic(base_size = 13) + 
  theme(plot.title = element_text(size=14)) +
  theme(axis.text = element_text(size = 14, color = "black"),
        axis.title.x = element_text(color="black", size=14),
        axis.title.y = element_text(color="black", size=14))
PboutWI

PboutSeas <- ggarrange(PboutSU,PboutWI,nrow=2)
PboutSeas

# Plot seasonal daytime mean state bouts
PboutDay <-stateBoutSummary %>% 
  group_by(season,ID,dayNight,timeBin,state) %>%
  summarize(mboutLengthMin=(mean(boutLengthSec)/60)) %>%
  ungroup %>% # Do I need to ungroup and regroup?
  group_by(season,ID,dayNight,timeBin,state,mboutLengthMin) %>%
  filter(dayNight=="day",
         state=="active") %>%
  ggplot() +
  geom_boxplot(aes(x = timeBin, y = mboutLengthMin, color = season)) +
  labs(x="Hour of Day",
       y = "Bout Length (min)",
       title = "Seasonal daytime mean active state by solar hour") + 
  theme_classic(base_size = 13) + 
  theme(plot.title = element_text(size=14)) +
  theme(axis.text = element_text(size = 14, color = "black"),
        axis.title.x = element_text(color="black", size=14),
        axis.title.y = element_text(color="black", size=14))
PboutDay

```

### Plot distribution of daytime state bouts (min) by state and season

```{r warning=FALSE}
# Plot distribution of daytime state bouts by austral season
distboutLengthDay <- stateBoutSummary %>% 
  group_by(season,state_classif) %>%
  filter(dayNight=="day") %>% 
  ggplot()+
  geom_histogram(aes(boutLength))+
  labs(x="State Bouts (min)",
       y = "Frequency",
       title = "Distribution of daytime state bouts by austral season") + 
  facet_grid(season~state_classif)
distboutLengthDay
ggsave(distboutLengthDay, file="Distribution of daytime state bouts by austral season.png",width=12, height=6, dpi=300)

# Plot distribution of nighttime state bouts by austral season
distboutLengthNight <- stateBoutSummary %>% 
  group_by(season,state_classif) %>%
  filter(dayNight=="night") %>% 
  ggplot()+
  geom_histogram(aes(boutLength))+
  labs(x="State Bouts (min)",
       y = "Frequency",
       title = "Distribution of nighttime state bouts by austral season") + 
  facet_grid(season~state_classif)
distboutLengthNight
ggsave(distboutLengthNight, file="Distribution of nighttime state bouts by austral season.png",width=12, height=6, dpi=300)

# Plot distribution of daytime state 4 bouts per bird by austral season
distboutLengthDay4 <- stateBoutSummary %>% 
  group_by(ID,season,state_classif) %>%
  filter(dayNight=="day",
         state_classif=="4") %>% 
  ggplot()+
  geom_histogram(aes(boutLength))+
  labs(x="State Bouts (min)",
       y = "Frequency",
       title = "Distribution of daytime state 4 bouts per bird by austral season") + 
  facet_grid(season~ID)
distboutLengthDay4
ggsave(distboutLengthDay4, file="Distribution of daytime state 4 bouts per bird by austral season.png",width=12, height=6, dpi=300)

# Plot distribution of daytime state 3 bouts per bird by austral season
distboutLengthDay3 <- stateBoutSummary %>% 
  group_by(ID,season,state_classif) %>%
  filter(dayNight=="day",
         state_classif=="3") %>% 
  # ggplot() + 
  #    geom_histogram(aes(boutLength)) + 
  #    facet_wrap(~season, scales = 'free', nrow = 2, strip.position = 'right')
  ggplot()+
  geom_histogram(aes(boutLength))+
  labs(x="State Bouts (min)",
       y = "Frequency",
       title = "Distribution of daytime state 3 bouts per bird by austral season") +
  facet_grid(season~ID)
distboutLengthDay3
ggsave(distboutLengthDay3, file="Distribution of daytime state 3 bouts per bird by austral season.png",width=12, height=6, dpi=300)

rm(distboutLengthDay,distboutLengthNight,distboutLengthDay3,distboutLengthDay4)

# State 3 attempt at histogram by individual
# scales = "free_x" not working
PhistState3 <- data_states %>%
  # group_by(ID,season,dayNight) %>% 
  filter(season=="Winter",
         dayNight == "day",
         state_classif =="3") %>%
  ggplot() +
  geom_bar(aes(x =state_classif,color = ID),fill="white",position="dodge") +
  # geom_vline()
  labs(x="State 3",
       y = "Frequency",
       title = "Austral winter daytime frequency of state 3") + 
  theme_classic(base_size = 13) + 
  theme(plot.title = element_text(size=14)) +
  theme(axis.text = element_text(size = 14, color = "black"),
        axis.title.x = element_text(color="black", size=14),
        axis.title.y = element_text(color="black", size=14))
PhistState3

# State 4 attempt at histogram by individual
PhistState4 <- data_states %>%
  # group_by(ID,season,dayNight) %>% 
  filter(season=="Winter",
         dayNight == "day",
         state_classif =="4") %>%
  ggplot() +
  geom_bar(aes(x =state_classif,color = ID),fill="white",position="dodge") +
  # geom_vline()
  labs(x="State 4",
       y = "Frequency",
       title = "Austral winter daytime frequency of state 4") + 
  theme_classic(base_size = 13) + 
  theme(plot.title = element_text(size=14)) +
  theme(axis.text = element_text(size = 14, color = "black"),
        axis.title.x = element_text(color="black", size=14),
        axis.title.y = element_text(color="black", size=14))
PhistState4

# Stack plots to compare full cycle
PhistState34 <- ggarrange(PhistState3,PhistState4,nrow=2)
# Save a file at 300 ppi
ggsave(PhistState34, file="Winter daytime frequencies of state 3 and state 4.png",width=8, height=8, dpi=300)
rm(PhistState3,PhistState4)
```

### Test to compare state proportions and absolute durations per period

```{r warning=FALSE}
# Test for difference in state duration between summer and winter day
# This subsets by day/night and filters for season, day, and state 4.
seasdiffdur1 <- stateSummaryIND2period %>% 
  subset(.,select = c(season,dayNight,state_classif,stateCountHour)) %>%
  filter(dayNight == "day",
         season == "Winter",
         state_classif == "4")
  # spread(season,stateCountHour)
seasdiffdur1$stateCountHourWI <- seasdiffdur1$stateCountHour
seasdiffdur1$stateCountHour <- NULL
seasdiffdur2 <- stateSummaryIND2period %>% 
  subset(.,select = c(season,dayNight,state_classif,stateCountHour)) %>%
  filter(dayNight == "day",
         season == "Summer",
         state_classif == "4")
  # spread(season,stateCountHour) 
seasdiffdur2$stateCountHourSU <- seasdiffdur2$stateCountHour
seasdiffdur2$stateCountHour <- NULL
shapiro.test(seasdiffdur1$stateCountHourWI)
shapiro.test(seasdiffdur2$stateCountHourSU)
test <- stats::t.test(seasdiffdur1$stateCountHourWI,seasdiffdur2$stateCountHourSU)
test

seasdiffdurState4 <- cbind(seasdiffdur1$stateCountHourWI,seasdiffdur2$stateCountHourSU)
colnames(seasdiffdurState4) <- c("stateCountHourWI", "stateCountHourSU")
# write_csv(as.data.frame(seasdiffdurState4),file.path(getwd(),"state4duration.csv"))
```

### Plots raw data distribution of states by season
*Irrelevant because variable of interest is ODBA daily sum not raw values. 
*This shows raw ODBA (1-Hz) is left skewed.
```{r warning=FALSE}
# Plot distribution of ODBA values
dist <- data_states %>% 
  ggplot()+
  geom_histogram(aes(ODBA))
dist

# Plot distribution of ODBA values by state (ALL INDIVIDUALS COMBINED)
dist_state <- data_states %>% 
  group_by(season,state_classif) %>%
  ggplot()+
  geom_histogram(aes(ODBA))+
  facet_grid(season~state_classif)
dist_state
ggsave(dist_state, file="ODBA distributions grouped by season and state.png",width=12, height=6, dpi=300)

data_states$yr <- year(data_states$dttz)
data_states$yr <- as.factor(data_states$yr)
data_states$ID <- as.factor(data_states$ID)
data_states$state_classif <- as.factor(data_states$state_classif)

```


### Calculate mean and SD for total ODBA over 24 hours.
*KJH: This needs work.
```{r warning=FALSE}
# Mean and SD of summer total ODBA per solar hour

# Mean and SD of winter total ODBA per solar hour
sumODBAWImean <- mean(sumODBAWI$sumODBA)
sumODBAWIsd <- sd(sumODBAWI$sumODBA)
summarySumODBA <- sumODBA %>% 
  group_by(season) %>% 
  summarize(mean(sumODBA),
            sd(sumODBA))
```

### Plot of seasonal mean oDBA per solar hour
This could be used to see if there are daily patterns in actity levels by hour across individuals.

```{r warning=FALSE}
# KJH: This needs work.

# Plot summer mean ODBA by solar hour
PodbaMeanSU <- data_states %>%
  filter(season == "Summer") %>% #View
  group_by(ID,timeBin) %>%
  summarize(odbaGMean = mean(ODBA/9.81,na.rm = TRUE)) %>% 
  ggplot() +
  geom_line(aes(x=timeBin, y=odbaGMean, group=ID,color=ID)) +
  labs(x="Solar Hour",
       y = "Mean ODBA (g)",
       title = "Austral summer mean ODBA per solar hour") + 
  theme_classic(base_size = 13) + 
  theme(plot.title = element_text(size=14)) +
  theme(axis.text = element_text(size = 14, color = "black"),
        axis.title.x = element_text(color="black", size=14),
        axis.title.y = element_text(color="black", size=14))
# PodbaMeanSU

# Plot winter mean ODBA by solar hour
PodbaMeanWI <- data_states %>%
  filter(season == "Winter") %>% #View
  group_by(ID, timeBin) %>%
  summarize(odbaGMean = mean(ODBA/9.81,na.rm = TRUE)) %>% 
  ggplot() +
  geom_line(aes(x=timeBin, y=odbaGMean, group=ID),alpha=0.7) +
  labs(x="Solar Hour",
       y = "Mean ODBA (g)",
       title = "Austral winter mean ODBA per solar hour") + 
  theme_classic(base_size = 13) + 
  theme(plot.title = element_text(size=14)) +
  theme(axis.text = element_text(size = 14, color = "black"),
        axis.title.x = element_text(color="black", size=14),
        axis.title.y = element_text(color="black", size=14))
# PodbaMeanWi

PodbaMeanSeas <- ggarrange(PodbaMeanSU,PodbaMeanWI,nrow=2)
PodbaMeanSeas

# Save a file at 300 ppi
# ggsave(PodbaMeanSeas, file="Seasonal mean ODBA per solar hour_black.png",width=12, height=6, dpi=300)
```

### Plot of seasonal daytime absolute duration of active state
Not including plot in thesis. Will include verbal results.
```{r warning=FALSE}
# Plot daytime absolute state durations
PdayDur <- stateSummaryIND2period2state %>%
  group_by(ID,season,dayNight) %>% 
  filter(dayNight == "day") %>%
  ggplot() +
  geom_boxplot(aes(x=state,y=stateCountHour, color = season)) +
  scale_color_discrete(name = "Season") +
  scale_y_continuous(breaks = seq(0, 24, by = 2)) +
  labs(x="State",
       y = "Duration (hr)",
       title = "Seasonal daytime state absolute durations") + 
  theme_classic(base_size = 13) + 
  theme(plot.title = element_text(size=14)) +
  theme(axis.text = element_text(size = 14, color = "black"),
        axis.title.x = element_text(color="black", size=14),
        axis.title.y = element_text(color="black", size=14))
PdayDur
# Save a file at 300 ppi
# ggsave(PdayDur, file="Seasonal daytime absolute state durations of active and rest.png",width=12, height=8, dpi=300)
```


### Plot seasonal nighttime activity absolute duration by solar hour
*Excluded nighttime period from analysis. 
```{r warning=FALSE}
# Plot summer nighttime activity absolute duration by hour
PcountSUnhr <- stateBout %>% 
  group_by(season,ID,dayNight,timeBin,state) %>%
  summarize(stateCount=n()/60) %>%
  filter(season=="Summer",
         dayNight=="night") %>%
  ggplot() +
  geom_boxplot(aes(x = timeBin, y = stateCount, 
                   color = state)) +
  labs(x="Solar Hour of Day",
       y = "Duration (min)",
       title = "Summer nighttime activity absolute duration by hour (n=12)") + 
  theme_classic(base_size = 13) + 
  theme(plot.title = element_text(size=14)) +
  theme(axis.text = element_text(size = 14, color = "black"),
        axis.title.x = element_text(color="black", size=14),
        axis.title.y = element_text(color="black", size=14))
# PcountSUnhr

# Plot winter nighttime activity absolute duration by hour
PcountWInhr <- stateBout %>% 
  group_by(season,ID,dayNight,timeBin,state) %>%
  summarize(stateCount=n()/60) %>%
  filter(season=="Winter",
         dayNight=="night") %>%
  ggplot() +
  geom_boxplot(aes(x = timeBin, y = stateCount, 
                   color = state)) +
  labs(x="Solar Hour of Day",
       y = "Duration (min)",
       title = "Winter nighttime activity absolute duration by hour (n=12)") + 
  theme_classic(base_size = 13) + 
  theme(plot.title = element_text(size=14)) +
  theme(axis.text = element_text(size = 14, color = "black"),
        axis.title.x = element_text(color="black", size=14),
        axis.title.y = element_text(color="black", size=14))
# PcountWInhr

PcountSeasnhr <- ggarrange(PcountSUnhr,PcountWInhr,nrow=2)
PcountSeasnhr
```

