---
title: "Analyze Rediscritized Acceleration Data"
author: "Katie Harrington"
date: "January 9, 2019"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float:
     collapsed: false
     smooth_scroll: false
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Get Started
*This script takes a CSV file with regularly sampled, calibrated acceleration data.

## Load libraries; set working directory and local time zone offset
```{r}
#Libraries utilized
library(tidyverse)
library(dplyr)
library(zoo)
library(tagtools)
library(lubridate)
library(plotly)
library(ggplot2)
library(RcppRoll)
library(car)
library(gridExtra)
library(signal)

setwd("~/Projects/R/TWlogger")
tzOffset <-"Etc/GMT+3"
```

# Prep Data
## Import Data: Option 1
*Use this import option for original 50-Hz data. 
*Includes downsampling script.
```{r DataImport, warning=FALSE, message=FALSE, collapse = TRUE}
# Select file to import
filename <- file.choose()
# Import original 50-Hz data
data <- read_csv(filename, 
                 col_types = cols(
                   #dttz = col_datetime(),
                   #dt = col_datetime(),
                   temp = col_double(), # Only comment this out for 2017/2019 tags
                   Ax = col_double(),
                   Ay = col_double(),
                   Az = col_double(),
                   Mx = col_double(),
                   My = col_double(),
                   Mz = col_double(),
                   # freq = col_double(), # Comment this out for 2017/2019 tags
                   secs_since = col_double()))

# Create deployment ID
depid <- basename(filename)
depid <- strsplit(depid,'-')
depid <- depid[[1]][1]
depid

# Confirm correct dttz, since it has not been retaining time zone when writing to CSV 
attr(data$dttz, "tzone") # Check tz
# Fix Option 1
# attr(data$dttz, "tzone") <- tzOffset # Use this to change from UTC to proper local time (time will change)
# Fix Option 2
data$dttz <- force_tz(data$dttz,tzone=tzOffset) # If time is correct and tz is wrong, force the tz (time will NOT change)
attr(data$dttz, "tzone") # Check tz

# Subset to only include time and acc data
data2 <- data[,c("dttz","true_since","Ax","Ay","Az")]

# Down sample data (decimate each vector separately)
df <- 50 # Set decimation factor df
fs <- 50 # Set original sampling rate

# For datetime select every nth value
dttz <- data2$dttz
a <- dttz
dttz_down <- a[seq(1, length(a), df)]

# For true_since select every nth value
true_since <- data2$true_since
a <- true_since
true_since_down <- a[seq(1, length(a), df)]

# Create individual vectors from acc fields
Ax <- data2$Ax
Ay <- data2$Ay
Az <- data2$Az

# Convert vectors to numeric matrix
Ax_mat <- matrix(Ax,ncol=1)
Ay_mat <- matrix(Ay,ncol=1)
Az_mat <- matrix(Az,ncol=1)

# Use decimate function
n <- 12*5
Ax_down <- decimate(Ax_mat,5,n=n,ftype="fir")
Ay_down <- decimate(Ay_mat,5,n=n,ftype="fir")
Az_down <- decimate(Az_mat,5,n=n,ftype="fir")

n <-12*10
Ax_down <- decimate(Ax_down,10,n=n,ftype="fir")
Ay_down <- decimate(Ay_down,10,n=n,ftype="fir")
Az_down <- decimate(Az_down,10,n=n,ftype="fir")

# Combine down sampled data into one dataframe
data2_down <- cbind.data.frame(dttz_down,true_since_down,Ax_down,Ay_down,Az_down)
data2 <- data2_down

# Create Bird ID after downsampling
depid2 <- strsplit(depid,'_')
depid2[[1]][3]
data2$ID <- depid2[[1]][3]

# Change column names to more practical shorter names
colnames(data2)[1:6] <- c("dttz","true_since","Ax","Ay","Az","ID")
# Reorder columns
data2 <- data2[,c("ID","dttz","true_since","Ax","Ay","Az")]

# Very important to run this line to reset sampling rate to down sampled rate
fs = fs/df

# Preserve original import data (data) and downsampled data (data2)
datax <- data2

# NOTE: Datax includes only down sampled acceleration data (IT HAS NOT BEEN SUBSET TO 24-hr)
# NOTE: Skip below Import Option 2 to subset and calculate metrics
```


## Import Data: Option 2
*Use this option to import previously down sampled 1-Hz data.
*Confirm this data already includes Bird ID and metrics.
```{r}
# Select file to import
filename <- file.choose()
# Import down sampled data
test <- read_csv(filename,
                 col_types = cols(
                   dttz_down = col_datetime(),
                   true_since_down = col_double(),
                   Ax_down = col_double(),
                   Ay_down = col_double(),
                   Az_down = col_double()))

# Create deployment ID
depid <- basename(filename)
depid <- strsplit(depid,'-')
depid <- depid[[1]][1]
depid

# Confirm correct dttz, since it has not been retaining time zone when writing to CSV 
attr(data$dttz, "tzone") # Check tz
# Fix Option 1
attr(data$dttz, "tzone") <- tzOffset # Use this to change from UTC to proper local time (time will change)
# Fix Option 2
data$dttz <- force_tz(data$dttz,tzone=tzOffset) # If time is correct and tz is wrong, force the tz (time will NOT change)
attr(data$dttz, "tzone") # Check tz

# Change column names to more practical shorter names
colnames(data)[1:9] <- c("ID","dttz","true_since","Ax","Ay","Az","Amag","Amag_rollmean","odba")

# Set sampling rate
# Need to write code so this calculates automatically, but haven't yet
fs = 1

# Preserve original import data (data)
datax <- data

# Datax includes down sampled acceleration data and metrics (IT HAS NOT BEEN SUBSET TO 24-hr)
```

## Subset data to timerange of interest
*Import Option 2 is already subset to 24-hr.
```{r}
# Confirm length of datax in hours
length(datax$dttz)/60/60

# If necessary, subset data
startTime <- as.POSIXct(strptime("2019-02-21 16:00:00",format="%Y-%m-%d %H:%M:%S"),tz=tzOffset)
endTime <- as.POSIXct(strptime("2019-02-22 16:00:00",format="%Y-%m-%d %H:%M:%S"),tz=tzOffset)
# Extract data for the selected timerange
datax <- subset(datax, datax$dttz >= startTime & datax$dttz <= endTime)
```


## Calculate Metrics
*Import Option 2 may already include metrics.
```{r metrics, echo=FALSE}
# Confirm sampling rate
fs
sw <- fs # Sets sampling window to be equal sampling rate

# Magnitude of acceleration
Amag <- sqrt(datax$Ax^2 + datax$Ay^2 + datax$Az^2)
datax$Amag <-Amag

# Check values
max(datax$Amag)
mean(datax$Amag)
min(datax$Amag)

# Running mean of magnitude of acceleration
# for 1Hz data, a 10 second window (+5, -5), window size is n+1 (this assumes an even windowSize)
windowSize = 11 
Amag_rollmean_10 <- roll_mean(datax$Amag,n=windowSize,fill=NA)
head(Amag_rollmean_10) # this gives (n-1)/2 NA's at the front
tail(Amag_rollmean_10) # and (n-1)/2 NA's at the end
# replace the NA's at the beginning with the first good value
Amag_rollmean_10[seq(1,(windowSize-1)/2)] <- Amag_rollmean_10[(windowSize-1)/2+1] # (windowSize-1)/2+1 is the first good value
# replace the NA's at the end with the last good value
Amag_rollmean_10[seq(length(Amag_rollmean_10)-(windowSize-1)/2+1,length(Amag_rollmean_10))] <- Amag_rollmean_10[length(Amag_rollmean_10)-(windowSize-1)/2]
datax$Amag_rollmean_10 <- Amag_rollmean_10

# Check values
max(datax$Amag_rollmean_10)
mean(datax$Amag_rollmean_10)
min(datax$Amag_rollmean_10)

windowSize =31 
Amag_rollmean_30 <- roll_mean(datax$Amag,n=windowSize,fill=NA)
head(Amag_rollmean_30) # this gives (n-1)/2 NA's at the front
tail(Amag_rollmean_30) # and (n-1)/2 NA's at the end
# replace the NA's at the beginning with the first good value
Amag_rollmean_30[seq(1,(windowSize-1)/2)] <- Amag_rollmean_30[(windowSize-1)/2+1] # (windowSize-1)/2+1 is the first good value
# replace the NA's at the end with the last good value
Amag_rollmean_30[seq(length(Amag_rollmean_30)-(windowSize-1)/2+1,length(Amag_rollmean_30))] <- Amag_rollmean_30[length(Amag_rollmean_30)-(windowSize-1)/2]
datax$Amag_rollmean_30 <- Amag_rollmean_30

# Check values
max(datax$Amag_rollmean_30)
mean(datax$Amag_rollmean_30)
min(datax$Amag_rollmean_30)

# Create matrix of acceleration data
A <- cbind(datax$Ax,datax$Ay,datax$Az)

# Calculate ODBA (Need to read about Wilson method, filter pass, and n)
A <- cbind(datax$Ax,datax$Ay,datax$Az)
datax$odba <- odba(A, sampling_rate = fs,method="wilson",n = sw) # n is sampling window, e.g. 50=1s
# Simple plot
# ba <- list(odba = odba)
# plott(ba, fs=fs) # NOTE: Change if different sampling rate

# Calculate jerk (sqrt of sum of squares of each axis)
datax$jerk <-njerk(A,sampling_rate=fs)
# Simple plot
# jlist <-list(jerk = jerk)
# plott(jlist,fs=fs)

# Check for NAs
sapply(datax, function(x) sum(is.na(x)))
# str(datax)

##########################################
####      Intellectual Excercise     #####
##########################################
# # Try to understand how Amag_rollmean relates to jerk
# same <- data$Amag_rollmean-data$jerk # How similar are the values? 
# slist <-list(same)
# plott(slist,1)
# # How do the means compare?
# Amag_m <- mean(data$Amag_rollmean)
# Amag_m
# jerk_m <-mean(data$jerk)
# jerk_m
# Amag_diff <-diff(data$Amag_rollmean) 
# Amag_diff_m <-mean(Amag_diff)
# Amag_diff_m

##########################################
####  Calculate Additional Metrics   #####
##########################################

# # Pitch and roll (NOTE: calculates in radians)
# pr <- a2pr(A,fs)
# #prlist <-list(pitch=pr$p,roll=pr$r)
# #plott(prlist,fs=fs)
# p <- pr$p
# r <-pr$r
# 
# # MSA (minimum specific acceleration)
# msa <-msa(A)
# #msalist <- list(msa = msa)
# #plott(msalist,fs=fs)
# 
# # Norm
# normAcc <- norm2(A)
# # normlist <- list(normAcc = normAcc)
# # plott(normlist,fs=fs)
# 
# # Running means and variances for each axis
# # fill=NA replaces missing values created by moving window (e.g.first 24 and last 25 rows are empty)
# # n is window size
# mheave <-roll_mean(data2$Az, n=sw, fill=NA)
# mheave[seq(1,24)] <- mheave[25]
# mheave[seq(length(mheave)-24,length(mheave))] <- mheave[length(mheave)-25]
# 
# varheave <-roll_var(data2$Az, n=sw, fill=NA)
# varheave[seq(1,24)] <- varheave[25]
# varheave[seq(length(varheave)-24,length(varheave))] <- varheave[length(varheave)-25]
# 
# msurge <-roll_mean(data2$Ax, n=sw, fill=NA)
# msurge[seq(1,24)] <- msurge[25]
# msurge[seq(length(msurge)-24,length(msurge))] <- msurge[length(msurge)-25]
# 
# varsurge <- roll_var(data2$Ax, n=sw, fill=NA)
# varsurge[seq(1,24)] <- varsurge[25]
# varsurge[seq(length(varsurge)-24,length(varsurge))] <- varsurge[length(varsurge)-25]
# 
# msway <- roll_mean(data2$Ay, n=sw, fill=NA)
# msway[seq(1,24)] <- msway[25]
# msway[seq(length(msway)-24,length(msway))] <- msway[length(msway)-25]
# 
# varsway <- roll_var(data2$Ay, n=sw, fill=NA)
# varsway[seq(1,24)] <- varsway[25]
# varsway[seq(length(varsway)-24,length(varsway))] <- varsway[length(varsway)-25]
```

## Plot Data
*Plots include: (1) full 24-hr acceleration data, (2) 30-min subsets of acceleration data, (3) and full 24-hr individual metrics.
```{r}
# Run this to create stacked plots 
# grid.arrange(p1, p2, p3, nrow=3)

# Plot acc
p1<- datax %>%
  gather(axis, acc, Ax:Az) %>%
  ggplot(aes(dttz, acc, color = axis)) +
  theme(legend.position="top") +
  geom_line() +
  theme_classic() +
  labs(x = "Time", y = "Acceleration")
p1

# Plot Amag_rollmean
p2 <- datax %>%
  ggplot(aes(dttz, y = Amag_rollmean_10),color = 'black') +
  geom_line() +
  theme_classic() +
  labs(x = "Time", y = "Running mean of magnitude of acceleration (10 sec window)")
p2

# Plot Amag_rollmean
p3 <- datax %>%
  ggplot(aes(x = dttz, y = Amag_rollmean_30),color= 'black') +
  geom_line() +
  labs(x = "Time", y = "Running mean of magnitude of acceleration (30 sec window)")
p3

# Plot ODBA
p4 <- datax %>%
  ggplot(aes(x = dttz, y = odba),color= 'black') +
  geom_line() +
  labs(x = "Time", y = "ODBA")
p4

# Plot jerk
p5<- datax %>%
  ggplot(aes(x = dttz, y = jerk),color= 'black') +
  geom_line() +
  labs(x = "Time", y = "Jerk")
p5

# Plot successive subsets of data and save to file
chunkLength <- 30 # Set length (in minutes) of chunks to plot
# i = 1 # Can set number of plots to print and run for loop from test to p
for(i in 1:ceiling(nrow(datax)/(chunkLength *60*50))) {
  test <- datax[seq(((i-1)*chunkLength*50*60)+1,i*chunkLength*50*60),]
  p <- test %>%
    gather(axis, acc, Ax:Az) %>%
    ggplot(aes(dttz, acc, color = axis)) +
    geom_line() +
    theme_classic() +
    labs(x = "Time", y = "Acceleration") +
    ggtitle(depid)
  p
  # save plots as .png
  ggsave(p, file=paste(depid,'-',i, ".png", sep=''), scale=2)
}

## Plot individual metrics and axes
## Select metric or axis to plot (do this by uncommenting selection)
# z <- msa
# z <- norm
# z <- mheave
# z <- msurge
# z <- mway
# z <- varheave
# z <- varsurge
# z <- varsway
# varlist <- list(z = z)
# plott(varlist,fs=fs)

## Plot multiple metrics
# multiple <- list(jerk = jerk,odba = odba,msa = msa)
# plott(multiple,fs=fs)
```

## Write to CSV
```{r}
write_csv(datax, file.path(dirname(filename), paste(depid,"-",fs,"Hz.csv",sep="")))
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
