---
title: "Seasonal Mass"
author: "Katie Harrington"
date: "June 20, 2019"
output: html_document
---

```{r setup, include=FALSE}
library(ggpubr) # needed for some plots
library(sjstats) # few useful packages
library(car) # for levenes test for equal variances
```

## Does body mass vary by time of year? 
Caracaras are sexually dimorphic (females larger than males). Resources vary among islands and prey availability (e.g. seabird success) could vary across years. Age could also predict mass fluctuations, since juveniles are less experienced at aquiring resources during their first winter.  

Description of data:   
Each bird only sampled once. Multiple samples per year day (either on same day specifically or on same year day of a different year). Not every year day has samples. Could use season as an independent variable, although that may mask the signal (e.g. if weight loss occurs at end of winter into early spring).  

Response variable:   
mass (continuous)   

Independent variables:   
year day (categorical 1-365) OR season (4 levels)  
age (4 levels), sex (2 levels)  
island (4 levels)  
year (10 levels)  

## Prepare data

### Import master banding file
This creates and saves bandMass.RData. Only need to do this once.

```{r}
# Import master banding file
band <- read.csv(file.choose(),na.strings=c("NA","NaN", " ",""))
names(band)
# Rename columns
colnames(band)[1:10] <- c("ID","color","date","age","mass","sex","origBand","island","loc","crop")
str(band$date)
# Create proper date object
band$date <-as.Date(band$date,format = "%m/%d/%Y")
# Add year day and season
band$yday <- yday(band$date)
band$season <- rep('NA')
# Calculated from 2018 solstices
band[which(band$yday > 79 & band$yday <= 173),]$season <- 'Fall'
band[which(band$yday > 173 & band$yday <= 265),]$season <- 'Winter'
band[which(band$yday > 265 & band$yday < 355),]$season <- 'Spring'
band[which(band$yday >= 355 | band$yday <= 79),]$season <- 'Summer'
band$season <- as.factor(band$season)
unique(band$season)
band$yr <- year(band$date)
# Try to remove all irregularities in character strings
unique(band$sex)
band$sex <- as.character(band$sex)
band$sex[band$sex == "Female"] <- "female"
band$sex[band$sex == "Male"] <- "male"
unique(band$sex)
band$sex <- as.factor(band$sex)

# Save dataframe for reimport later
save(band,file="bandMass.RData")

```

### Import recapture data
This creates and saves recapMass.RData. Only need to do this once.

```{r}
# Import recapture data
filename <- file.choose()
recap <- read_csv(filename, 
                 col_types = cols(),
                 na = c(" ", "ND"))
str(recap)
# Subset to only include rows with mass recorded
recap <- subset(recap, recap$Mass > 0)
# Remove unused columns
names(recap)
recap <-recap[,c(-2,-3,-5,-11:-15)]
names(recap)
# Rename columns
colnames(recap)[1:7] <- c("date","ID","age","mass","crop","island","loc")
# Create proper date object
str(recap$date)
recap$date <-as.Date(recap$date,format = "%m/%d/%Y")
# Create year day and season
recap$yday <- yday(recap$date)
recap$season <- rep('NA')
# Calculated from 2018 solstices
recap[which(recap$yday > 79 & recap$yday <= 173),]$season <- 'Fall'
recap[which(recap$yday > 173 & recap$yday <= 265),]$season <- 'Winter'
recap[which(recap$yday > 265 & recap$yday < 355),]$season <- 'Spring'
recap[which(recap$yday >= 355 | recap$yday <= 79),]$season <- 'Summer'
recap$season <- as.factor(recap$season)
unique(recap$season)

# Save dataframe for reimport later
save(recap,file="recapMass.RData")

```

## Explore data

```{r}
# Import R.data files
load("bandMass.RData")
load("recapMass.RData")
```

### Create summary tables

```{r}
# Create summary table for master band data
test <- band %>% 
  group_by(crop) %>% 
  summarize(count=n())

# All mass data (male and female combined)
# Need to update to omit NA from mass calculations
bandSum <- band %>%
  group_by(island,season,crop) %>%
  filter(crop == 0) %>% 
  summarize(mMass = mean(mass),
            sdMass = sd(mass),
            minMass = min(mass),
            maxMass = max(mass),
            count = n())

bandSumM <- band %>% 
  filter(sex=="male") %>%
  group_by(island,season,crop) %>% 
  filter(crop==0) %>% 
  summarize(mMass = mean(mass),
            sdMass = sd(mass),
            minMass = min(mass),
            maxMass = max(mass),
            count = n())

bandSumF <- band %>% 
  filter(sex=="female") %>% 
  group_by(island,season,crop) %>% 
  filter(crop==0) %>% 
  summarize(mMass = mean(mass),
            sdMass = sd(mass),
            minMass = min(mass),
            maxMass = max(mass),
            count = n())

# Create summary table of mean mass of birds with zero crop
recapSum <- recap %>% 
  group_by(island,season,crop) %>% 
  filter(crop == 0) %>% 
  summarize(mMass = mean(mass),
            sdMass = sd(mass),
            minMass = min(mass),
            maxMass = max(mass),
            count = n())

# Create summary table for TWLogger data

```


### Density distribution of all mass data (islands and age pooled)

```{r}
# Male and female combined
dMass <- density(band$mass,na.rm = TRUE)
plot(dMass)

# Only male
dMassM <- density(bandM$mass,na.rm = TRUE)
plot(dMassM)

# Only female
dMassF <- density(bandF$mass,na.rm=TRUE)
plot(dMassF)

```


### Test seasonal differences on SDI
```{r}
SDImassM <- band %>%
  filter(sex=="male",
         island=="SDI",
         age=="JUV",
         crop==0)
SDImassF <- band %>%
  filter(sex=="female",
         island=="SDI",
         age=="JUV",
         crop==0)

SDImassMsum <- band %>%
  filter(sex=="male",
         island=="SDI",
         crop==0) %>%
  group_by(season,age) %>% 
  summarize(count = n(),
            mMass = mean(mass),
            sdMass = sd(mass),
            minMass = min(mass),
            maxMass = max(mass))

SDImassFsum <- band %>%
  filter(sex=="female",
         island=="SDI",
         crop==0) %>% 
  group_by(season,age) %>% 
  summarize(count=n(),
            mMass = mean(mass),
            sdMass = sd(mass),
            minMass = min(mass),
            maxMass = max(mass))
  
# JUVENILE MALES -- test for seasonal differences
shapiro.test(SDImassM$mass) # Results: p-value = 0.01418
var.test(mass ~ season,data=SDImassM) # Results: p-value = 0.503
# Data is nonparametric and of EQUAL variance
wilcox.test(mass ~ season,data=SDImassF) # Results: W = 3319, p-value = 9.243e-06

# JUVENILE FEMALES -- test for seasonal differences
shapiro.test(SDImassF$mass) # Results: p-value = 0.02413
var.test(mass ~ season,data=SDImassF) # Results: p-value = 0.005141
# Data is nonparametric and of unequal variance
wilcox.test(mass ~ season,data=SDImassF) # Results: W = 545.5, p-value = 0.04365





# Test for seasonal differences among males (with age as covariate)
result1 <- aov(mass~season + age,data = SDImassM) # Results: both season and age influence mass on Saunders. 
summary(result1)

result2 <- aov(mass~season * age,data = SDImassM) # Results: the effect of season does not depend on age.
summary(result2)

anova(result1,result2) #Results: models are not significantly different.

TukeyHSD(result1)

# Test for seasonal differences among females (with age as covariate)
result3 <- aov(mass~season + age,data = SDImassF) # Results:
summary(result3)

result4 <- aov(mass~season * age,data = SDImassF) # Results:
summary(result4)

anova(result3,result4) #Results: models are not significantly different.

dMass <- density(band$mass,na.rm = TRUE)
plot(dMass)

bandM <- band %>% 
  filter(sex=="Male")
bandF <- band %>% 
  filter(sex=="FeMale")


#KJH: begin here by running post hoc and plotting residuals


```




```{r}
# To do: ANOVA and density distribution by yday. By grouping am I missing the signal?

bandM <- band %>% 
  filter(sex=="Male")
bandF <- band %>% 
  filter(sex=="FeMale")

dMassM <- density(bandM$mass,na.rm = TRUE)
plot(dMassM)
dMassF <- density(bandF$mass,na.rm=TRUE)
plot(dMassF)


# Test for differences among males
# ggplot(data=bandM, aes(x=season, y =mass))+geom_point(aes(colour = island))
# par(mfrow=c(1,2))
# plot(mass ~ season + island, data=bandM)
# m2 = aov(mass ~ season + yr, data = bandM)
# summary(m2)
# TukeyHSD(m2)

# Second attempt
library(lme4)
glmer(y~Probe + Extraction + Dilution + (1|Tank), family=binomial, data=mydata)

# Third attempt 
#H0(mass): μ1 = μ2 = μ3 = μ4
#Ha(mass): μ1 != μ2 != μ3 != μ4
#Run one-way ANOVA
m1 = lm(mass ~ season,data = bandM)
anova(m1)
qqnorm(m1$res)
plot(m1$fitted,m1$res,xlab="Fitted",ylab="Residuals")
#Test for equal variances. Bartlett if data normal, levenes better if data not normal.
bartlett.test(mass ~ season, data = bandM) #if p>0.05, then variances are equal
leveneTest(mass ~ season, data = bandM)
#Results: We have insufficient evidence to conclude that mass differs among seasons (F = 0.99, df = 3, p = 0.40).
#Run ANCOVA with no interation
#Assumptions: independent observations, 2 predictors (one categorical, one continuous), response is continuous, residuals are normally distributed, equal variances
m2 = lm(mass ~ season + island, data = bandM)
anova(m2)
#Results: Island influences seasonal mass (F3,3 = 24.07, p < 0.001), though season does not (F3,3 = 1.11, p = 0.34).
summary(m2)
#Examine residuals
plot(m2)
shapiro.test(m2$residuals)
#Run ANCOVA with interactions
m3 = lm(mass ~ season * island, data = bandM)
anova(m3)

# KJH still need to modify all of this:

#Results: After controlling for variability in islands, there is a difference in mass (F3,11 = 129.2,  p < 0.01). There is no interaction (F2,9 = 0.12, p = 0.89) between genotype and temperature on metabolic rate.
#Assess the assumptions
plot(m3) #plots a few graphs to assess residuals
shapiro.test(m3$residuals) #performs normality test on residuals
#Because interaction was not significant, remove and rerun the model without it (same as part A). Because this is the better model, we will perform post hoc tests on teh ANCOVA without interactions (m2).
m2a = aov(m2) #Create an anova summary using aov in order to perform post hoc tests
TukeyHSD(m2a) #Tukey's is the most straight forward post hoc test for ANCOVAs in R.
#Results: Genotype aa has a higher metabolic rate than the other genotpes (Tukey's HSD posthoc, p < 0.05).
#Results: Genotype aa differs from AA and from Aa. There is no difference between AA and Aa.
ggqqplot(m2a$residuals)
ggdensity(m2a$residuals)
shapiro.test(m2a$residuals) #Residuals are normal and look okay, so they meet assumptions.
#Results: Genotype and temperature influence fish metabolic rate (ANCOVA, Genotype F2,11 = 18.65, p = 0.0003; Temp F1,11 = 572.03, p < 0.0001). Genotype aa has a higher metabolic rate than genotypes AA and Aa (Tukey's HSD < 0.05). Metabolic rate is higher at higher temperatures.

```

Redux
```{r}

SDImass <- band %>%
  filter(island=="SDI")
SDImassF <- band %>%
  filter(island=="SDI",
         sex=="FeMale")
SDImassM <- band %>%
  filter(island=="SDI",
         sex=="Male")

dSDImassF <- density(SDImassF$mass,na.rm = TRUE)
plot(dSDImassF)
dSDImassM <- density(SDImassM$mass,na.rm = TRUE)
plot(dSDImassM)

shapiro.test(SDImassF$mass)
shapiro.test(SDImassM$mass)

m1 <- aov(mass ~ season + age, data = SDImassF)
summary(m1)
m2 <- aov(mass ~ season * age, data = SDImassF)
summary(m2)
TukeyHSD(m1)
```


### Seasonal body mass: TW Logger deployments

```{r}
# Select file to import
filename <- file.choose()
# Import original 50-Hz data
meta <- read_csv(filename, 
                 col_types = cols())
meta
names(meta)
colnames(meta)[1:16] <- c("type","version","model","tagID","tagWeight","ID","age","sex","mass","deployDate","deployLoc","retrieveDate","retrieveLoc","tot","data","tagPercent")
meta$notes <- NULL
# Remove unused rows
meta <-meta[,c(-2,-3,-4,-5,-12,-13,-14,-16,-17)]
metaTW <- filter(meta,type !="GPS")
metaTW$sex <- NULL
# Rename ID to remove band color
ID <-strsplit(as.character(metaTW$ID), " ")
ID <- sapply(ID, "[[", 1)
metaTW1 <- metaTW %>% 
  separate(ID,c("ID","color"))
metaTW$ID <- metaTW1[,2]
metaTW$season <- rep('NA')
# metaTW[which(metaTW$deployLoc == "Neck")]$season <- "Summer" # Couldn't get this to work
metaTW$type <- c(rep("V1",12),rep("V2",15))
metaTW$season <-c(rep("Summer",6),rep("Winter",14),rep("Summer",7))
metaTWsum <- metaTW %>%
  group_by(season) %>% 
  summarize(mMass = mean(mass),
            sdMass = sd(mass),
            minMass = min(mass),
            maxMass = max(mass)) %>% View


# Test for normality
shapiro.test(metaTW$mass) # Results: p-value = 0.181
# Test for equal variances
var.test(mass ~ deployLoc,data=metaTW) # Results: p-value = 0.6912
# Test for seasonal mass difference 
t.test(mass ~ deployLoc, data=metaTW,var.equal = TRUE) # Results: t = -0.069918, df = 25, p-value = 0.9448

# Save TWLogger mass by season
save(metaTW,file="TWLogger_mass.RData")
```


### Plot master banding data

```{r}

# Meanline summer
meanlineMassM <- band %>%
  filter(sex == "Male",
         crop == 0) %>%
  group_by(yday) %>%
  summarize(mMassM = mean(mass))

shadeMass = data.frame(x1=c(0,173,355),
                     x2=c(79,265,365),
                     y1=c(0,0,0),
                     y2=c(2100,2100,2100))

pMassM <- band %>%
  filter(sex=="Male",
         crop == 0) %>% #View
  ggplot() +
  geom_boxplot(aes(x=season, y=mass, color=island),alpha=0.7) +
  scale_y_continuous(breaks = seq(0,2100, by = 100), expand = c(0, 0)) +
  scale_x_discrete() +
  theme_classic(base_size = 13)
  
pMassM


```


